---
title: "Data manipulation practice: Sunshine list 2019"
output: html_document
---

## Importing libraries
```{r}
library(dplyr)
library(ggplot2)

options(scipen = 20)
```

## Reading Sunshine List 2019 data set, basic data cleaning
```{r}
# importing Ontario 2019 Sunshine list
dfRaw <- read.csv("SunshineList2019.csv", stringsAsFactors = FALSE)

# checking duplicates, none as equal number of unique and total rows
nrow(unique(dfRaw))
nrow(dfRaw)
# making secondary df that will be modified in later sections
dfCleaned <- dfRaw
# cleaning Salary.Paid col, removing $ and , then transforming col into numeric
dfCleaned$Salary.Paid <- as.numeric(gsub('[$,]', '', dfCleaned$Salary.Paid))
```


## Basic exploration, number of different unique sectors, positions, distribution of salaries
```{r}
# num of unique sectors is 17
unique(dfRaw$Sector)
# num of unique employers is 1898
length(unique(dfRaw$Employer))
# num of unique job titles is 32868
length(unique(dfRaw$Job.Title))


# graphing distribution for all salaries
dfGraphAllSalaries <- dfCleaned
ggplot(data = dfGraphAllSalaries, aes(Salary.Paid)) + geom_histogram(binwidth=1000)

# graphing distribution for salaries > 250k
dfGraph250kSalaries <- dfGraphAllSalaries[!(dfGraphAllSalaries$Salary.Paid < 250000), ]
ggplot(data = dfGraph250kSalaries, aes(Salary.Paid)) + geom_histogram(binwidth=1000)

```




## Listing total salary by sector
```{r}
# summarizing df by sum of Salary.Paid by sector
dfSector <- dfCleaned
dfSector <- group_by(dfSector, Sector)
dfSector <- summarize(dfSector, total = sum(Salary.Paid))

# graphing total salaries by sector
ggplot(dfSector, aes(x = Sector, y = total)) + 
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))

# graph in log10 to show categories such as seconded that are too low to appear
ggplot(dfSector, aes(x = Sector, y = log10(total))) + 
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))
```


## Institutions by number of people in the list
```{r}
# df of number of employed by employer
dfNumInstit <- dfCleaned %>% group_by(Employer) %>% summarize(count = n()) %>% arrange(desc(count))
dfNumInstit

# graphing overall distribution
ggplot(dfNumInstit, aes(Employer, count)) +
  geom_point() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```


## People by their salaries and their institutions
```{r}
# df of people and their respective institution listed by descending order of salary
dfTopPeople <- dfCleaned %>% group_by(Last.Name, First.Name, Employer) %>% 
  arrange(desc(Salary.Paid))
dfTopPeople
```


## Scatterplot of top 25 titles in average comp and their number of people in the list
```{r, fig.width = 12}
# df of mean salary by job title and number of people of that title
dfMeanCompNumTitles <- dfCleaned %>% group_by(Job.Title) %>% 
  summarize(mean = mean(Salary.Paid), count = n()) %>% arrange(desc(mean))
dfMeanCompNumTitles

# graph of top 25 job titles in pay compensation matching the count threshold of 20
dfCompTitlesTop25 <- dfMeanCompNumTitles %>% filter(count > 19) %>% head(25)
ggplot(dfCompTitlesTop25, aes(x = Job.Title, y = mean)) +
  geom_point() +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))

# graph of bottom 25 job titles in pay compensation matching the count threshold of 20
dfCompTitlesBot25 <- dfMeanCompNumTitles %>% filter(count > 19) %>% tail(25)
ggplot(dfCompTitlesBot25, aes(x = Job.Title, y = mean)) +
  geom_point() +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))


# graph of top 50 most frequent job titles
dfFreqTitlesTop50 <- dfMeanCompNumTitles %>% arrange(desc(count)) %>% head(50)
ggplot(dfFreqTitlesTop50, aes(x = Job.Title, y = count)) +
  geom_point() +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))

# respective avg salaries for the top 50 most freq job titles
dfFreqTitlesTop50Comp <- dfMeanCompNumTitles %>% filter(Job.Title %in% dfFreqTitlesTop50$Job.Title)
ggplot(dfFreqTitlesTop50Comp, aes(x = Job.Title, y = mean)) +
  geom_point() +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))
```

